\section{The Spec() functor}
We aim to mirror the previous considerations of \cref{mot-circle} in the case of an arbitrary ring.

\begin{df}
Let $A$ be a ring. The \textit{spectrum} of $A$ is its set of prime ideals:
\[\spec(A) = \{ \mathfrak p \text{ a prime ideal in } A \}.\]
\end{df}

\begin{example}
  $\spec(\complex [x]) =  \{0\} \cup \{ (xf-a) \suchthat a \in \complex \}$.
\end{example}

For a ring homomorphism $f \colon A \to B$, \cref{primes are preserved under preimage} asserts that there is a map
\[f^* \colon \spec(B) \to \spec(A)\]
given by 
\[B \supseteq \mathfrak q \mapsto f^{-1}(\mathfrak q) \subseteq A.\]

We will now move towards upgrading spectra of rings (which up to this point we considered as mere sets) to topological spaces in such a way that makes the maps $f^*$ continuous. In this, we follow the previous motivation.

\begin{df}
For $E \subseteq A$ an arbitrary subset, we define the \textit{vanishing locus of $E$} as the set
\[V(E) = \{ \mathfrak p \in \spec(A) \suchthat E \subseteq \mathfrak p \} .\]
\end{df}

\begin{prop}
  \label{zariski-prop}
  The sets $V(E)$, $E \subseteq A$, are the closed subsets of a topology.
\end{prop}

\begin{proof}
  One sees immediately that
  \[ V(1) = \emptyset, \quad V(0) = \spec(A).\]
  Elementary set-theoretic considerations reveal that
  \[\cap_{i \in I} V(E_i) = V(\cup_{i \in I} E_i).\]

  It now suffices to show that
\[ V(E_1) \cup V(E_2) = V(E_1 \cdot E_2), \quad E_1 \cdot E_2 \coloneqq \{ e_1 \cdot e_2 \suchthat e_1 \in E_1, e_2 \in E_2 \}.\]

The inclusion $V(E_1) \cup V(E_2) \subseteq V(E_1 \cdot E_2)$ is easy to verify. Suppose without loss of generality that $\mathfrak p \in V(E_1)$, that is $\mathfrak p \supseteq E_1$. Then
\[ E_1 \cdot E_2 \subseteq \mathfrak p \cdot E_2 \subseteq \mathfrak p,\]
because $\mathfrak p$ is an ideal.

In order to prove the other inclusion, we need to use that $\mathfrak p$ is prime. Suppose that
\[ \mathfrak p \not\supseteq E_1 \quad \text{and} \quad \mathfrak p \not\supseteq E_2.\]
Then there exist $e_i \in E_i \setminus \mathfrak p$, $i=1,2$. For those,
\( e_1 \cdot e_2 \in E_1 \cdot E_2 \setminus \mathfrak p\)
holds because $\mathfrak p$ is prime.
\end{proof}

\begin{df}
  The topology on $\spec(A)$ defined by \cref{zariski-prop} is called the \textit{Zariski topology}.
\end{df}

\begin{question}
What is the Zariski topology for $A = \complex [x]$?
\end{question}

\begin{answer}
We note that
\begin{align*}
  V(f_1, \ldots, f_r) & = \{ \mathfrak p \text{ prime in } A \suchthat f_1, \dotsc, f_r \in \mathfrak p\} \\
                      & = \{(x-a) \text{ maximal in } A \suchthat f_1, \dotsc, f_r \in (x-a)\} \\
                      & = \{ (x-a) \suchthat f_1(a) = \dotsc = f_r(a) = 0 \}.
\end{align*}

If any $f_i$ is nonzero, then $0 \notin V(f_1, \dotsc, f_r)$.

Hence,
\[ V(f_1, \ldots, f_r) = \text{ set of common roots of }f_1, \ldots, f_r \]
and one sees that $V(f_1, \ldots, f_r)$ is finite.

In fact, all finite subsets of the set of maximal ideals are of the form $V(f_1, \ldots, f_r)$. Note that the only closed set containing the prime ideal $0$ is $\spec(A)$ itself; in other words, $0$ lies in every nonempty open set. This yields the cofinite topology augmented by $\{0\}$.
\end{answer}

\begin{note}
  Because the closure of $\{0\} \subseteq \spec(A)$ is $\spec(A)$ itself, the resulting space is not T1, let alone Hausdorff. It is, however, T0.
\end{note}

\begin{prop}
  Let $f \colon A \to B$ be a ring homomorphism. Then the induced map
  \[f^* \colon \spec(B) \to \spec(A)
    \quad \text{with} \quad
    \mathfrak q \mapsto f^{-1}(\mathfrak q)\]
  is continuous.
\end{prop}
\begin{proof}
It suffices to see that the preimage of every closed set is closed. Indeed, we claim that
\[ (f^*)^{-1}(V(E)) = V(f(E)).\]
We unravel the definitions:
\begin{align*}
  (f^*)^{-1}(V(E)) & = \{ \mathfrak p \subseteq B \suchthat f^*(\mathfrak p) \in V(E)\} \\
                   & = \{ \mathfrak p \subseteq B \suchthat f^{-1}(\mathfrak p) \in V(E)\} \\
                   & = \{ \mathfrak p \subseteq B \suchthat E \subseteq f^{-1}(\mathfrak p) \} \\
                   & = \{ \mathfrak p \subseteq B \suchthat f(E) \subseteq \mathfrak p\} \\
                   & = V(f(E)).
\end{align*}
Note that the considerations here are purely set-theoretical. The condition that $f$ be a ring homomorphism is only relevant for $f^*$ to be well defined on spectra.
\end{proof}

The upshot is: when $k = \bar k$, $\spec_{max}(k[x_1, \ldots, x_n])$ is well-behaved. We will see later, in the Nullstellensatz, that all maximal ideals in $k[x_1, \ldots, x_n]$ are of the form $(x_1 - a_1, \ldots x_n - a_n)$.

% In fact,
% $\spec_{max}(k[x_1, \ldots, x_n]) \cong k^n$
% under
% $(x_1 - a_1, \ldots, x_n - a_n) \mapsto (a_1, \ldots, a_n)$.

% The intuition here is: $k[x_1, \ldots, x_n]$ is the ring of regular functions on $k^n$.
% TODO explain or remove

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../commalg"
%%% End:
